---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rppo

The following is an R package for accessing the plant phenology ontology (PPO) global data store and terms contained in the
Plant Phenology Ontology. For more information visit the [PPO global data portal](http://plantphenology.org/) or the 
[ppo-data-pipeline git repository](https://github.com/biocodellc/ppo-data-pipeline)
## Installation

You can install rppo from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("biocodellc/rppo")
```

## Example

Query the plant phenology ontology and return a list of present or absent stages
you can use the stageIDs returned from this function to query stages in the ppo_data function

```{r}
present_stages <- ppo_terms(present=T)
absent_stages <- ppo_terms(absent=T)
```

A simple query example to populate a data frame
```{r}
df <- ppo_data(genus = "Quercus", fromYear = 2013, toYear = 2013, fromDay = 100, toDay = 110,termID='obo:PPO_0002313')
```

An example showing how we can use the results data frame from the ppo_data function in conjunction with other tools
```{r}
require("leaflet")        

# run a query to obtain observations
df <- ppo_data(bbox='44,-124,46,-122', fromDay = 1, toDay = 60)

# add the data to leaflet... NOTE: the following is to quickly visualize a small number (<10,000 records)
# adding more markers to a map than this could cause problems
map <- leaflet(df)
# addMarkers automatically looks at latitude/longitude column names from dataframe
map = addMarkers(map)
map = addProviderTiles(map, "Esri.NatGeoWorldMap")
map
```

## Citation
To cite package 'rppo' in publications use:

```
   'John Deck' (2018). rppo: an interface to the Plant Phenology Ontology and associated data store.  R package version 0.1
   https://github.com/biocodellc/rppo
```
