---
output: github_document
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rppo
[![Build Status](https://travis-ci.org/biocodellc/rppo.svg?branch=master)](https://travis-ci.org/biocodellc/rppo)
[![codecov.io](https://codecov.io/github/r-lib/covr/coverage.svg?branch=master)](https://codecov.io/github/r-lib/covr?branch=master)
[![](https://badges.ropensci.org/175_status.svg)](https://github.com/ropensci/onboarding/issues/207)




The global plant phenology data portal, or [PPO data portal](https://www.plantphenology.org/), is an aggregation of 
plant phenological observations from [USA-NPN](https://www.usanpn.org/usa-national-phenology-network), [NEON](https://www.neonscience.org/), and [PEP725](http://www.pep725.eu/) representing 
20 million phenological observations from across North America and Europe.  The PPO data portal utilizes the [Plant
Phenology Ontology](https://github.com/PlantPhenoOntology/ppo/) (PPO) to align phenological
terms and measurements from the various databases. The rppo R package enables programmatic access to all data contained in the PPO data portal incuding selected classes contained in the PPO itself.  

For more information visit how data is assembled for the PPO data portal, visit the 
[ppo-data-pipeline git repository](https://github.com/biocodellc/ppo-data-pipeline).

## Installation

The production version of rppo can be installed with (when available on CRAN):
```{r}
#install.packages("rppo")  # this will work once package is available on CRAN
#library(rppo)
```

You can install the development version of rppo from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("biocodellc/rppo")
```


Install the library once it has been downloaded from CRAN or github.

```{r install library}
library(rppo)
```

## Examples

We reccomend visiting the [rppo vignette](http://htmlpreview.github.io/?https://github.com/biocodellc/rppo/blob/master/vignettes/rppo-vignette.html) for a set of well-explained examples on using the rppo package, as well as viewing man pages for rppo functions in the R environment, using `?ppo_data` and `?ppo_terms`.  Following are a couple of brief examples to illustrate how to get started:

```{r examples}
# query all results from day 1 through 100 in a particular bounding box, 
# limited to 2 records
r <- ppo_data(fromDay = 1, toDay = 100, bbox="37,-120,38,-119", limit=2)

# view the data returned
print(r$data)

# view the number of possible records returned
print(r$number_possible)

# return a data frame of present
presentTerms <- ppo_terms(present = TRUE)

# print the 2nd present term returned
print(presentTerms[2,])
```

## Citation
To cite the 'rppo' R package in publications use:

```
   'John Deck, Brian Stucky, Ramona Walls, Kjell Bolmgren, Ellen Denny, Robert Guralnick' (2018). rppo: An interface to the Plant Phenology Ontology and associated data store.  R package version 1.0
   https://github.com/biocodellc/rppo
```

## Code of Conduct
View our [code of conduct](CONDUCT.md)
