---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rppo
[![Build Status](https://travis-ci.org/biocodellc/rppo.svg?branch=master)](https://travis-ci.org/biocodellc/rppo)


The global plant phenology data portal (PPO data portal) is an aggregation of 
plant phenological observations from [USA-NPN](https://www.usanpn.org/usa-national-phenology-network), [NEON](https://www.neonscience.org/), and [PEP725](http://www.pep725.eu/) representing 
20 million phenological observations from across North America and Europe.  The PPO data portal utilizes the [Plant
Phenology Ontology](https://github.com/PlantPhenoOntology/ppo/) to align phenological
terms and measurements from the various databases. The rppo R package enables programmatic access to all data contained in the PPO data portal. 

For more information visit the [PPO global data portal](http://plantphenology.org/) or the 
[ppo-data-pipeline git repository](https://github.com/biocodellc/ppo-data-pipeline).

## Installation

The production version of rppo can be installed with (when available on CRAN):
```{r}
#install.packages("rppo")
#library(rppo)
```

You can install the development version of rppo from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("biocodellc/rppo")
```



## Example

Query the plant phenology ontology and return a list of present or absent terms.
Use the termIDs returned from this function to query terms in the ppo_data function

```{r term example}
library(rppo)
present_terms <- ppo_terms(present=T)
```

A simple query example. The results of the ppo_data function are returned as a list.  Hence, to retrieve the data portion of the returned results you will need to access the "data" element of the returned list.
```{r data example}
results <- ppo_data(genus = "Quercus", fromYear = 2013, toYear = 2013, fromDay = 100, toDay = 110,termID='obo:PPO_0002313', limit=10)
df <- results$data
readme <- results$readme
citation <- results$citation

# use cat(readme) and cat(citation) to view the text output
```


## Citation
To cite the 'rppo' R package in publications use:

```
   'John Deck, Brian Stucky, Ramona Walls, Kjell Bolmgren, Ellen Denny, Robert Guralnick' (2018). rppo: An interface to the Plant Phenology Ontology and associated data store.  R package version 1.0
   https://github.com/biocodellc/rppo
```

## Code of Conduct
View our [code of conduct](CONDUCT.md)
