---
title: "rppo vignette"
author: "John Deck"
date: "`r Sys.Date()`"
output:
 html_document:
    keep_md: yes
vignette: |
  %\VignetteIndexEntry{rppo Vignette} 
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "rppo-vignette-"
)
library(rppo)
```

The rppo package contains just two functions.  One to query terms from the Plant Phenology Ontology (PPO) and another to query the data global plant phenology data portal (PPO data portal).  Following are three examples which illustrate use of these functions: the first two sections illustrate the `ppo_data` and `ppo_terms` functions and the third section illustrates how to use the functions together.

### ppo_terms function
It is frequently useful to look through the list of present and absent terms contained in the PPO.   The `ppo_terms` function returns present terms, absent terms, or both, with columns containing a termID, label, definition and full URI for each term.  Use the termIDs returned from this function to query terms in the `ppo_data` function.  The following example returns the present terms into a "present_terms" data frame and a sample slice from the dataframe.

```{r term example, echo=TRUE, message=FALSE, warning=FALSE}
present_terms <- ppo_terms(present = TRUE)
# print the first five rows, with just the termIDs and labels
print(present_terms[1:5,c("termID","label")])
```

### ppo_data function
The `ppo_data` function queries the PPO Data Portal, passing values to the database and extracting matching results. The results of the `ppo_data` function are returned as a list with five elements: 1) a data frame containing data, 2) a readme string containing usage information and some statistics about the query itself, 3) a citation string containing information about proper citation, 4) a number_possible integer indicating the total number of results if a limit has been specified, and 5) a status code returned from the service. 

The "df" variable below is populated with results from the data element in the results list, with an example slice of data showing the first record.
```{r data example, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
results <- ppo_data(genus = "Quercus", fromYear = 2013, toYear = 2013, fromDay = 100, toDay = 110, termID = 'obo:PPO_0002313', limit = 10, timeLimit = 2)
df <- results$data
print(df[1:1,])
```

The readme and citation files returned by the list of results can be accessed by calling the readme and citation elements.  Note that the the file "citation_and_data_use_policies.txt" that is referred to in the readme file can be accessed using `cat(results$citation)` 
```{r readme results example}
cat(results$readme)
```

The results lists also shows the number of possible results in the results set, which is useful if the submitted query had a limit.  For example, in the query above, the limit is set to 10 but we want to know how many records were possible if the limit was not set.
```{r readme possible results example}
cat(results$number_possible)
```

### working with terms and data together
Here we will generate a data frame showing the frequency of "present" and "absent" terms for a particular query.  The query is for genus = "Quercus" and latitude > 47.  For each row in the returned data frame `ppo_data` will typically return multiple terms in the termID field, corresponding to phenological stages as defined by the PPO.  For our example, we will generate a frequency table of the number of times "present" or "absent" term occur in the entire returned dataset.  Note that the termID field returned by `ppo_data` will return "presence" terms in addition to "present" and "absent" terms, while the `ppo_terms` function only returns "present" and "absent" terms.  Thus, our frequency distribution only counts the number of "present" and "absent" terms [For an in-depth discussion of the difference between "presence", "present", and "absent", see https://www.frontiersin.org/articles/10.3389/fpls.2018.00517/full].  Finally, since termIDs are returned as URI identifiers and not easily readable text, this example maps termIDs to labels. The resulting data frame shows two columns: 1) a column of term labels, and 2) a frequency of the number of times this label appeared in the result set. 

```{r workting with terms and data example}
###############################################################################
# Generate a frequency data frame showing the number of times each termID
# is populated for genus equals "Quercus" above latitude of 47
# Note that all latitude/longitude queries need to be in the format of a
# bounding box
###############################################################################
df <- ppo_data(
  genus = "Quercus", 
  bbox="47,-180,90,180", timeLimit = 4)
# return just the termID column
t1 <- df$data[,c('termID')]
# paste each cell into one string
t2<-paste(t1, collapse = ",")
# split strings at ,
t3<-strsplit(t2, ",")
# create a frequency table as a data frame
freqFrame <- as.data.frame(table(t3))

# create a new data frame that we want to populate
resultFrame <- data.frame(
  label = character(), 
  frequency = integer(), 
  stringsAsFactors = FALSE)

###############################################################################
# Replace termIDs with labels in frequency frame
###############################################################################
# fetch "present" and "absent" terms using `ppo_terms`
termList <- ppo_terms(absent = TRUE, present = TRUE, timeLimit = 2);

# loop all "present"" and "absent" terms
if (!is.null(termList)) {
  for (term in 1:nrow(termList)) {
    termListTermID<-termList[term,'termID'];
    termListLabel<-termList[term,'label'];
    # loop all rows that have a frequency generated
    for (row in 1:nrow(freqFrame)) {
      freqFrameTermID = freqFrame[row,'t3']
      freqFrameFrequency = freqFrame[row,'Freq']
      # Populate resultFrame with matching "present" or "absent" labels.
      # In this step, we will ignore "presence" terms
      # found in the frequency frame since the ppo_terms only returns
      # "present" and "absent" terms. 
      if (freqFrameTermID == termListTermID) {
        resultFrame[nrow(resultFrame)+1,] <- c(termListLabel,freqFrameFrequency)
      }
    }
  }
} else {
  message("termList is null, likely due to a server response issue.  Try increasing the timeLimit or try again later. If the problem persists email the authors.")
}


# print results, showing term labels and a frequency count
print(resultFrame)
```
